---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output:   
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one dataset, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)

```{r}
# Loading packages
library(pacman)
p_load(tidyverse, tidymodels, groupdata2)
```

## Creating training and test data set
```{r}
# Reading data
data <- read_csv("data_AF.csv")

# Removing NA's
df <- filter(data, !is.na(ScaledPitchIQR))
df <- filter(df, !is.na(ScaledSpokenProp))
df <- filter(df, !is.na(ScaledSpeechRate))
df <- filter(df, !is.na(ScaledPauseDur))

#Changing Diagnosis to factor
df$Diagnosis <- as.factor(df$Diagnosis)

# Partitioning the data
set.seed(69)
df_list <- partition(df, p = 0.2, cat_col = c("Diagnosis"), id_col = c("uniquepairID"), list_out = T)
df_test = df_list[[1]]
df_train = df_list[[2]]


ls.str(df_train)
```


# Create recipe for ONLY ScaledSpeechRate as it was the feature that was described the best by Diagnosis
```{r}
rec <- df_train %>% recipe(Diagnosis ~ ScaledSpeechRate + ScaledPitchIQR + ScaledPauseDur) %>% # defines the outcome
  step_center(all_numeric()) %>% # center numeric predictors
 # step_scale(all_numeric()) %>% # scales numeric predictors
  step_corr(all_numeric()) %>% 
  check_missing(everything()) %>%
  prep(training = df_train)

train_baked <- juice(rec) # extract df_train from rec

rec #inspect rec

# Applying recipe to test
test_baked <- rec %>% bake(df_test)

# Seeing if our recipes are the same
juice(rec)
rec %>% bake(df_train) # the recipes are the same, however columns are changes

# Creating models
# logistic regression
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(Diagnosis ~ ScaledSpeechRate, data = train_baked)

# Applying model to test data
#predict class
log_class <- log_fit %>%
  predict(new_data = test_baked)
#get prob of class
log_prop <- log_fit %>%
  predict(new_data = test_baked, type = "prob") %>%
  pull(.pred_1)

#get multiple at once
test_results <- 
  test_baked %>% 
  select(Diagnosis) %>% 
  mutate(
    log_class = predict(log_fit, new_data = test_baked) %>% 
      pull(.pred_class),
    log_prob  = predict(log_fit, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1))

test_results %>% 
  head(5) 

# Performance metrics including PPV, NPV, sensitivity, specificity, kappa and accuracy
#full_eval <- as_tibble(bind_rows(ppv(test_results, truth = Diagnosis, estimate = log_class), 
#           npv(test_results, truth = Diagnosis, estimate = log_class),
#            sens(test_results, truth = Diagnosis, estimate = log_class),
#            spec(test_results, truth = Diagnosis, estimate = log_class),
#            metrics(test_results, truth = Diagnosis, estimate = log_class)))

multimetric <- metric_set(accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
full_eval <- multimetric(bind_cols(test_baked, test_results), truth = Diagnosis, estimate = log_class)  
full_eval

# plotting the roc curve:
test_results %>%
  roc_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()

# plotting the gain curve
test_results %>% 
  mutate(log_prob = 1 - log_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()
  
```

# Cross-Validation on ScaledSpeechRate
```{r}
p_load(gganimate, gifski, png)
# Multiple cross validation
set.seed(69)
# create 10 folds, 10 times, and make sure Suicide_Risk is balances across groups
cv_folds <- vfold_cv(df_train, v = 10, repeats = 10, strata = Diagnosis, group = uniquepairID)

#prepare data set and fetch train data
cv_folds <- cv_folds %>% 
  mutate(recipes = splits %>%
           # prepper is a wrapper for `prep()` which handles `split` objects
           map(prepper, recipe = rec),
         train_data = splits %>% map(training))

# train model of each fold
  # create a non-fitted model
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") 


cv_folds <- cv_folds %>%  mutate(
  log_fits = pmap(list(recipes, train_data), #input 
                            ~ fit(log_fit, formula(.x), data = bake(object = .x, new_data = .y)) # function to apply
                 ))
cv_folds %>% head(5)

predict_log <- function(split, rec, model) {
  # IN
    # split: a split data
    # rec: recipe to prepare the data
    # 
  # OUT
    # a tibble of the actual and predicted results
  baked_test <- bake(rec, testing(split))
  tibble(
    actual = baked_test$Diagnosis,
    predicted = predict(model, new_data = baked_test) %>% pull(.pred_class),
    prop_SCZ =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_1),
    prop_control =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_0)
  ) 
}

# apply our function to each split, which their respective recipes and models (in this case log fits) and save it to a new col
cv_folds <- cv_folds %>% 
  mutate(pred = pmap(list(splits, recipes, log_fits) , predict_log))

# Performance metrics
eval <- 
  cv_folds %>% 
  mutate(
    metrics = pmap(list(pred), ~ multimetric(., truth = actual, estimate = predicted, prop_SCZ))) %>% 
  select(id, id2, metrics) %>% 
  unnest(metrics)

#inspect performance metrics
multimetric2 <- metric_set(roc_auc, accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) %>% 
  arrange(desc(kap)) %>% 
  head()


# Creating dataset with mean of all metrics.
evalmetrics <- eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) 

mean_eval <- evalmetrics %>% summarize(
  mean(accuracy),
  mean(bal_accuracy),
  mean(kap),
  mean(npv),
  mean(ppv),
  mean(sens),
  mean(spec)
)
mean_eval %>% head()

cv_folds %>% 
  mutate(roc = pmap(list(pred), roc_curve, truth = actual, prop_SCZ)) %>% 
  unnest(roc) %>%
  mutate(idval=as.integer(gsub("Fold", "", id2))) %>% 
  ggplot() + transition_reveal(idval) + transition_reveal(sensitivity) +
  aes(x = 1 - specificity, y = sensitivity, color = id2) +
  geom_path() +
  geom_abline(lty = 3) + facet_wrap(~id)

anim_save("cv.png")
```

# Create recipe for ALL predictors to see if they are better at classifying
```{r}
rec <- df_train %>% recipe(Diagnosis ~ ScaledSpeechRate + ScaledPitchIQR + ScaledPauseDur) %>% # defines the outcome
  step_center(all_numeric()) %>% # center numeric predictors
 # step_scale(all_numeric()) %>% # scales numeric predictors
  step_corr(all_numeric()) %>% 
  check_missing(everything()) %>%
  prep(training = df_train)

train_baked <- juice(rec) # extract df_train from rec

rec #inspect rec

# Applying recipe to test
test_baked <- rec %>% bake(df_test)

# Seeing if our recipes are the same
juice(rec)
rec %>% bake(df_train) # the recipes are the same, however columns are changes

# Creating models
# logistic regression
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(Diagnosis ~., data = train_baked)

# Applying model to test data
#predict class
log_class <- log_fit %>%
  predict(new_data = test_baked)
#get prob of class
log_prop <- log_fit %>%
  predict(new_data = test_baked, type = "prob") %>%
  pull(.pred_1)

#get multiple at once
test_results <- 
  test_baked %>% 
  select(Diagnosis) %>% 
  mutate(
    log_class = predict(log_fit, new_data = test_baked) %>% 
      pull(.pred_class),
    log_prob  = predict(log_fit, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1))

test_results %>% 
  head(5) 

# Performance metrics including PPV, NPV, sensitivity, specificity, kappa and accuracy
#full_eval <- as_tibble(bind_rows(ppv(test_results, truth = Diagnosis, estimate = log_class), 
#           npv(test_results, truth = Diagnosis, estimate = log_class),
#            sens(test_results, truth = Diagnosis, estimate = log_class),
#            spec(test_results, truth = Diagnosis, estimate = log_class),
#            metrics(test_results, truth = Diagnosis, estimate = log_class)))

multimetric <- metric_set(accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
full_eval <- multimetric(bind_cols(test_baked, test_results), truth = Diagnosis, estimate = log_class)  
full_eval

# plotting the roc curve:
test_results %>%
  roc_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()

# plotting the gain curve
test_results %>% 
  mutate(log_prob = 1 - log_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()
```

# Cross-Validation on ALL predictors
```{r}
# Multiple cross validation
set.seed(69)
# create 10 folds, 10 times, and make sure Suicide_Risk is balances across groups
cv_folds <- vfold_cv(df_train, v = 10, repeats = 10, strata = Diagnosis, group = uniquepairID)

#prepare data set and fetch train data
cv_folds <- cv_folds %>% 
  mutate(recipes = splits %>%
           # prepper is a wrapper for `prep()` which handles `split` objects
           map(prepper, recipe = rec),
         train_data = splits %>% map(training))

# train model of each fold
  # create a non-fitted model
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") 


cv_folds <- cv_folds %>%  mutate(
  log_fits = pmap(list(recipes, train_data), #input 
                            ~ fit(log_fit, formula(.x), data = bake(object = .x, new_data = .y)) # function to apply
                 ))
cv_folds %>% head(5)

predict_log <- function(split, rec, model) {
  # IN
    # split: a split data
    # rec: recipe to prepare the data
    # 
  # OUT
    # a tibble of the actual and predicted results
  baked_test <- bake(rec, testing(split))
  tibble(
    actual = baked_test$Diagnosis,
    predicted = predict(model, new_data = baked_test) %>% pull(.pred_class),
    prop_SCZ =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_1),
    prop_control =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_0)
  ) 
}

# apply our function to each split, which their respective recipes and models (in this case log fits) and save it to a new col
cv_folds <- cv_folds %>% 
  mutate(pred = pmap(list(splits, recipes, log_fits) , predict_log))

# Performance metrics
eval <- 
  cv_folds %>% 
  mutate(
    metrics = pmap(list(pred), ~ multimetric(., truth = actual, estimate = predicted, prop_SCZ))) %>% 
  select(id, id2, metrics) %>% 
  unnest(metrics)

#inspect performance metrics
multimetric2 <- metric_set(roc_auc, accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) %>% 
  arrange(desc(kap)) %>% 
  head()


# Creating dataset with mean of all metrics.
evalmetrics <- eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) 

mean_eval <- evalmetrics %>% summarize(
  mean(accuracy),
  mean(bal_accuracy),
  mean(kap),
  mean(npv),
  mean(ppv),
  mean(sens),
  mean(spec)
)
mean_eval %>% head()

cv_folds %>% 
  mutate(roc = pmap(list(pred), roc_curve, truth = actual, prop_SCZ)) %>% 
  unnest(roc) %>% 
  ggplot() +
  aes(x = 1 - specificity, y = sensitivity, color = id2) +
  geom_path() +
  geom_abline(lty = 3) + facet_wrap(~id)
```


# Running model and evaluations for ScaledSpeechRate and ScaledPitchIQR
```{r}
rec <- df_train %>% recipe(Diagnosis ~ ScaledSpeechRate + ScaledPitchIQR) %>% # defines the outcome
  step_center(all_numeric()) %>% # center numeric predictors
 # step_scale(all_numeric()) %>% # scales numeric predictors
  step_corr(all_numeric()) %>% 
  check_missing(everything()) %>%
  prep(training = df_train)

train_baked <- juice(rec) # extract df_train from rec

rec #inspect rec

# Applying recipe to test
test_baked <- rec %>% bake(df_test)

# Seeing if our recipes are the same
juice(rec)
rec %>% bake(df_train) # the recipes are the same, however columns are changes

# Creating models
# logistic regression
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(Diagnosis ~ ScaledSpeechRate + ScaledPitchIQR, data = train_baked)

# Applying model to test data
#predict class
log_class <- log_fit %>%
  predict(new_data = test_baked)
#get prob of class
log_prop <- log_fit %>%
  predict(new_data = test_baked, type = "prob") %>%
  pull(.pred_1)

#get multiple at once
test_results <- 
  test_baked %>% 
  select(Diagnosis) %>% 
  mutate(
    log_class = predict(log_fit, new_data = test_baked) %>% 
      pull(.pred_class),
    log_prob  = predict(log_fit, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1))

test_results %>% 
  head(5) 

# Performance metrics including PPV, NPV, sensitivity, specificity, kappa and accuracy
#full_eval <- as_tibble(bind_rows(ppv(test_results, truth = Diagnosis, estimate = log_class), 
#           npv(test_results, truth = Diagnosis, estimate = log_class),
#            sens(test_results, truth = Diagnosis, estimate = log_class),
#            spec(test_results, truth = Diagnosis, estimate = log_class),
#            metrics(test_results, truth = Diagnosis, estimate = log_class)))

multimetric <- metric_set(accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
full_eval <- multimetric(bind_cols(test_baked, test_results), truth = Diagnosis, estimate = log_class)  
full_eval

# plotting the roc curve:
test_results %>%
  roc_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()

# plotting the gain curve
test_results %>% 
  mutate(log_prob = 1 - log_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = Diagnosis, log_prob) %>% 
  autoplot()
```


# Cross-Validation on two predictors
```{r}
# Multiple cross validation
set.seed(69)
# create 10 folds, 10 times, and make sure Suicide_Risk is balances across groups
cv_folds <- vfold_cv(df_train, v = 10, repeats = 10, strata = Diagnosis, group = uniquepairID)

#prepare data set and fetch train data
cv_folds <- cv_folds %>% 
  mutate(recipes = splits %>%
           # prepper is a wrapper for `prep()` which handles `split` objects
           map(prepper, recipe = rec),
         train_data = splits %>% map(training))

# train model of each fold
  # create a non-fitted model
log_fit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") 


cv_folds <- cv_folds %>%  mutate(
  log_fits = pmap(list(recipes, train_data), #input 
                            ~ fit(log_fit, formula(.x), data = bake(object = .x, new_data = .y)) # function to apply
                 ))
cv_folds %>% head(5)

predict_log <- function(split, rec, model) {
  # IN
    # split: a split data
    # rec: recipe to prepare the data
    # 
  # OUT
    # a tibble of the actual and predicted results
  baked_test <- bake(rec, testing(split))
  tibble(
    actual = baked_test$Diagnosis,
    predicted = predict(model, new_data = baked_test) %>% pull(.pred_class),
    prop_SCZ =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_1),
    prop_control =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_0)
  ) 
}

# apply our function to each split, which their respective recipes and models (in this case log fits) and save it to a new col
cv_folds <- cv_folds %>% 
  mutate(pred = pmap(list(splits, recipes, log_fits) , predict_log))

# Performance metrics
eval <- 
  cv_folds %>% 
  mutate(
    metrics = pmap(list(pred), ~ multimetric(., truth = actual, estimate = predicted, prop_SCZ))) %>% 
  select(id, id2, metrics) %>% 
  unnest(metrics)

#inspect performance metrics
multimetric2 <- metric_set(roc_auc, accuracy, bal_accuracy, sens, yardstick::spec, ppv, npv, kap)
eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) %>% 
  arrange(desc(kap)) %>% 
  head()


# Creating dataset with mean of all metrics.
evalmetrics <- eval %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) 

mean_eval <- evalmetrics %>% summarize(
  mean(accuracy),
  mean(bal_accuracy),
  mean(kap),
  mean(npv),
  mean(ppv),
  mean(sens),
  mean(spec)
)
mean_eval %>% head()

cv_folds %>% 
  mutate(roc = pmap(list(pred), roc_curve, truth = actual, prop_SCZ)) %>% 
  unnest(roc) %>% 
  ggplot() +
  aes(x = 1 - specificity, y = sensitivity, color = id2) +
  geom_path() +
  geom_abline(lty = 3) + facet_wrap(~id)
```


